```{r libraries}
options(scipen=999)

# Importing the libraries

library(tidyverse)
library(dplyr)
library(glue)
library(ggplot2)
library(ggcorrplot)
library(broom)
library(lubridate)
library(viridis)
library(scales)
```

Let's import the spotify data set.

```{r dataset}
# Importing the data set

tracks <- read_csv("https://raw.githubusercontent.com/sushmaakoju/spotify-tracks-data-analysis/main/SpotifyFeatures.csv")

```

Let's check for missing values in our data set.

```{r missingvalues}
#checking for missing values in the data set

colSums(is.na(tracks))
```

As per above summary, there are no missing values in this data set.

```{r distinctvalues}
# As there are no missing values, we will check for other errors.

# Let's check for distinct genre names to see if any genre is repeated.

distinct(data.frame(tracks$genre)) 
```

As we can see there are two Children's Music genre. We should rectify this typo so that a single genre is visible.

```{r cleandata}

# Rectifying typo in genre name Children's Music.

colnames(tracks)[colnames(tracks$genre) == "Children's Music"] <- "Children’s Music"

# Rechecking the distinct values.

distinct(data.frame(tracks$genre))
```

```{r mutate}
#Mutate variables from numeric to factor

tracks <- tracks  %>% 
  mutate(genre = as.factor(genre),
         key = as.factor(key),
         genre = as.factor(str_replace_all(genre, "[[:punct:]]", "")),
         mode = as.factor(mode))

summary(tracks)
```

```{r filterdata}

# Filter columns that are not presently needed for our analysis

tracks1 <- tracks %>% select(-c(track_id,time_signature))

# Glimpse data

glimpse(tracks1)

```

Now that we have pre-processed our data set, we can visualize basic data to find correlations among our variables.

From our data set, we can group our data set by genre and focus on the popularity parameter for visualization of data.

```{r plot_averagepopularity_genre}
genre_popularity <- tracks1 %>% select(popularity, genre) %>% group_by(genre) %>% summarise("average_popularity" = round(mean(popularity)))

p <- ggplot(data=genre_popularity, mapping = aes(x = reorder(genre,average_popularity), y = average_popularity, fill = genre)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  theme(
    legend.position = "none",
    
  ) +
  labs(
    y = "Average popularity",
    x = "Genre"
  )

p
```
We have plotted the music genre from highest popularity to lowest popularity.

We can see that the Top 5 popular genres on spotify are -
Pop 
Rap
Rock
HipHop
Dance

Genres with lowest popularity on spotify.

Anime, 
Comedy, 
Opera, 
Movie, and 
A Capella

Let's analyse the relationship between all numeric spotify features now using a correlation map.

```{r plotcorr}
# Plotting a correlation heat map to check correlation between all audio features.

ggcorr(tracks1, low = "blue", high = "red")
```

From above graph we can list out below significant observations -

1) Energy and loudness have the highest correlation.
2) Energy and acousticness have a highly-correlated inverse relationship
3) Loudness and acousticness have a highly-correlated inverse relationship
4) Some of the variables most correlated to Popularity seem to be: acousticness, danceability, energy, instrumentalness, and loudness.